---
title: "Exercise 2"
output: html_document
date: '2022-05-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Task 1: Import Data & Install Libraries

```{r}
## Load the necessary libraries ################################################

library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times
library(tidyverse)

## Import the downloaded csv ##################################################

wildschwein_BE <- read_delim("wildschwein_BE_2056.csv",",") # adjust path

wildschwein_BE

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)

wildschwein_BE$timelag <- as.numeric(difftime(lead(wildschwein_BE$DatetimeUTC), wildschwein_BE$DatetimeUTC, units = "secs"))

wildschwein_BE
```

Task 2: Getting an overview

How many individuals were tracked?

```{r}

unique(wildschwein_BE$TierID)
unique(wildschwein_BE$TierName)

```

...3 Individuals

Visual approach:

For how long were the individuals tracked? Are there gaps?

```{r}

wildschwein_dates <- wildschwein_BE %>%
  group_by(TierID) %>%
  summarise(
    startTime = min(DatetimeUTC),
    endTime = max(DatetimeUTC),
    duration = as.factor(as.integer(difftime(endTime, startTime, units = "days")))
  )
wildschwein_dates

wildschwein_dates %>%
  ggplot(aes(color = duration)) +
  geom_segment(aes(x = startTime, xend = endTime, y = TierID, yend = TierID), size = 0.5) +
  xlab("DatetimeUTC") +
  labs(color='Measurement \nduration [days]') 

```

Were all individuals tracked concurrently or sequentially?
Interpretation: All the animals were measured concurrently at least at point

What is the temporal sampling interval between the locations? => What does that mean?

```{r}

wildschwein_BE %>%
  ggplot() +
  geom_freqpoly(mapping = aes(x = timelag, colour = TierName)) +
  xlim(c(700, 1000)) +
  labs(color='Name of animal') 

```

Interpretation: I show that the large majority of timelags (for all three individuals) was somewhere between 850 and 950 seconds

Task 3: Deriving movement parameters I: Speed

```{r}

wildschwein_BE$steplength <- 
  ((lead(wildschwein_BE$E, 1) - wildschwein_BE$E)^2 + (lead(wildschwein_BE$N, 1) - wildschwein_BE$N)^2)^(1/2)
wildschwein_BE

wildschwein_BE$speed <- wildschwein_BE$steplength / wildschwein_BE$timelag
wildschwein_BE

```

What speed unit do you get?

Speed = Steplength per second => But what is steplength? => Apparently it's meter (see Task 4, last plot)
As such, speed = m/s

Task 4: Cross-scale movement analysis

```{r}

caro <- read_delim("caro60.csv")
caro
caro <- st_as_sf(caro, coords = c("E", "N"), crs = 2056, remove = FALSE)
caro

```

Calculating speed at different intervals:

```{r}

caro_3 <- caro %>%
  filter(row_number() == 1 | row_number() %% 3 == 0)

caro_6 <- caro %>%
  filter(row_number() == 1 | row_number() %% 6 == 0)

caro_9 <- caro %>%
  filter(row_number() == 1 | row_number() %% 9 == 0)

nrow(caro)
nrow(caro_3)
nrow(caro_6)
nrow(caro_9)

lead(caro$DatetimeUTC, 1)

```

(Here's a little function that does it for you)

```{r}
calc_step_and_speed <- function(df) {
  
  df <- mutate(df, timelag = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC, unit = "secs")))
  df <- mutate(df, steplength = ((lead(E, 1) - E)^2 + (lead(N, 1) - N)^2)^(1/2))
  df <- mutate(df, speed = steplength / timelag)
  
  return(df)
}

caro <- calc_step_and_speed(caro)
caro

caro_3 <- calc_step_and_speed(caro_3)
caro_3

caro_6 <- calc_step_and_speed(caro_6)
caro_6

caro_9 <- calc_step_and_speed(caro_9)
caro_9
```

Visualizing derived speed at different sampling intervals:

```{r}
caro_combined <- bind_rows("1 minute" = caro, "3 minute" = caro_3, "6 minute" = caro_6, "9 minute" = caro_9, .id = "groups")

caro_combined <- drop_na(caro_combined, speed)

caro_combined %>%
  ggplot(mapping = aes(x = DatetimeUTC, y = speed, colour = groups)) +
  geom_line() +
  labs(color='Measurement interval') 

```

Interpretation: The longer the intervals, the slower the derived speeds and the smaller the values of acceleration and deceleration. In other words: When deriving the sped at a low interval (i.e. 9 minutes), everything becomes a little calmer, the individuals are not speeding up / slowing down as quickly as when deriving the speed at a high interval (i.e. 1 minute)

Visualizing the trajectories:

```{r}
caro_combined %>%
  ggplot() +
  geom_path(mapping = aes(x = E, y = N)) +
  facet_wrap(~ groups, ncol = 2)
```

Interpretation: This graph is not as pretty as the example. However, it shows quickly that the interval at which speed is derived, has a strong effect on the mapping of the trajectories. Namely, the higher the interval (i.e. 1 minute), the rounder the edges and the more precise the trajectory. The lower the interval (i.e. 9 minutes), the sharper the edges and the rougher and less accurate the trajectories.

One more pretty graph:

```{r}
caro_combined %>%
  filter(groups == "1 minute" | groups == "3 minute") %>%
  ggplot(mapping = aes(x = E, y = N, color = groups)) +
  geom_point() +
  geom_path() +
  labs(color='Measurement Interval') 
```

Task 5: Deriving movement parameters II: Rolling window functions

```{r}

library(zoo)

example <- rnorm(10)
example

rollmean(example, k = 3, fill = NA, align = "left")
rollmean(example, k = 4, fill = NA, align = "left")

```

Plotting and comparing speed with different moving window intervalls:

```{r}
caro$rollmean_3 <- rollmean(caro$speed, k = 3, fill = NA, align = "left")
rm3_plot <- caro %>% 
  drop_na(rollmean_3) %>%
  ggplot(mapping = aes(x = DatetimeUTC, y = rollmean_3)) +
  geom_line() +
  labs(y = "Speed", x = "Time", title = "k = 3")
rm3_plot

caro$rollmean_4 <- rollmean(caro$speed, k = 4, fill = NA, align = "left")
rm4_plot <- caro %>% 
  drop_na(rollmean_4) %>%
  ggplot(mapping = aes(x = DatetimeUTC, y = rollmean_4)) +
  geom_line() +
  labs(y = "Speed", x = "Time", title = "k = 4")
rm4_plot

caro$rollmean_8 <- rollmean(caro$speed, k = 8, fill = NA, align = "left")
rm8_plot <- caro %>% 
  drop_na(rollmean_8) %>%
  ggplot(mapping = aes(x = DatetimeUTC, y = rollmean_8)) +
  geom_line() +
  labs(y = "Speed", x = "Time", title = "k = 8")
rm8_plot

orig_plot <- caro %>% 
  drop_na(speed) %>%
  ggplot(mapping = aes(x = DatetimeUTC, y = speed)) +
  geom_line() +
  labs(y = "Speed", x = "Time", title = "original")
orig_plot

```

Showing the derived speeds with different rolling windows all at once:

```{r}
library(cowplot)
plot_grid(orig_plot, rm3_plot, rm4_plot, rm8_plot)
```

Interpretation: The plot shows that the higher the rollmean value (k) (i.e. the larger the moving window), the rounder the curve becomes.
The edges become less sharp and we get a smoother curve